---
title: "Exploring Gapminder"
output: github_document
---

# Table of content
- Determining attributes of gapminder (Smell test the data) [Here](#Header_1)
- Exploring individual variables [Here](#Header_2)


<a name="Header_1">
</a>

# Determining attributes of gapminder (Smell test the data)


Loading the libraries

```{r,message=FALSE}
library(gapminder)
library(tidyverse)
library(ggplot2)
```

Q1) What type is gapminder?
```{r}
typeof(gapminder)
```

Q2) What is the class on gapminder

```{r}
class(gapminder)
```


Q3) How many variables/columns?

```{r}
ncol(gapminder)
```

Q4) How many rows/observations?

```{r}
nrow(gapminder)
```

Q5) Can you get these facts about “extent” or “size” in more than one way? Can you imagine different functions being useful in different contexts?
 
```{r}
#way 1 using dim command
dim(gapminder)[1] #Rows
dim(gapminder)[2] #Coloumns
#way two using length and transpose command
length(transpose(gapminder)) #Rows
length(gapminder)  #Coloumns
```

Q6) What data type is each variable?

```{r}
sapply(gapminder,typeof)
```

<a name="Header_2">
</a>

# Exploring individual variables

Q1) What are possible values (or range, whichever is appropriate) of each variable?

```{r}
max <- sapply(select(gapminder,year,lifeExp,pop,gdpPercap),max)
min <- sapply(select(gapminder,year,lifeExp,pop,gdpPercap),min)
#Range of Years
c(max[1],min[1])
#Range of LifeExp
c(max[2],min[2])
#Range of Pop
c(max[3],min[3])
#Range of gdpPercap
c(max[4],min[4])
table(gapminder$country)
table(gapminder$continent)
```

Another method using function range (simple!!!)

```{r}
sapply(select(gapminder,year,lifeExp,pop,gdpPercap),range)
```

Another method using summary (Yay!!!)

```{r}
summary(gapminder)
```

Q2) What values are typical? What’s the spread? What’s the distribution?

```{r}
ggplot(gapminder, aes(x = log10(gdpPercap), y = lifeExp)) +
  geom_point(aes(color = continent))
ggplot(gapminder, aes(x = continent, y = lifeExp)) + geom_jitter()+geom_boxplot()

```


```{r}
gapminder %>% 
  filter(country == "Canada") %>% 
  ggplot(aes(x = year, y =lifeExp))+geom_point()
```




