---
title: "Exploring Gapminder"
output: github_document
---

# Table of content
- Determining attributes of gapminder (Smell test the data) [Here](#Header_1)
      + What type is gapminder? [Here](#Subheader 1_1)
      + What is the class on gapminder? [Here](#Subheader 1_2)
      + How many variables/columns? [Here](#Subheader 1_3)
      + How many rows/observations? [Here](#Subheader 1_4)
      + Can you get these facts about “extent” or “size” in more than one way? Can you imagine different functions being useful in different contexts? [Here](#Subheader 1_5)
      + What data type is each variable? [Here](#Subheader 1_5)
  
- Exploring individual variables [Here](#Header_2)
      + What are possible values (or range, whichever is appropriate) of each variable?[Here](#Subheader 2_1) 
      + What values are typical? What’s the spread? What’s the distribution?[Here](#Subheader 2_2)
  
- Exploring various plot types [Here](#Header_3)
      + A scatterplot of two quantitative variables [Here](#Subheader 3_1)
      + A plot of one quantitative variable. Maybe a histogram or densityplot or frequency polygon [Here](#Subheader 3_2)
      + A plot of one quantitative variable and one categorical. Maybe boxplots for several continents or countries [Here](#Subheader 3_3)

<a name="Header_1">
</a>

# Determining attributes of gapminder (Smell test the data)


Loading the libraries

```{r,message=FALSE}
library(gapminder)
library(tidyverse)
library(ggplot2)
```

<a name="Subheader 1_1">
</a>


Q1) What type is gapminder?
```{r}
typeof(gapminder)
```


<a name="Subheader 1_2">
</a>

Q2) What is the class on gapminder

```{r}
class(gapminder)
```


<a name="Subheader 1_3">
</a>

Q3) How many variables/columns?

```{r}
ncol(gapminder)
```


<a name="Subheader 1_4">
</a>

Q4) How many rows/observations?

```{r}
nrow(gapminder)
```

<a name="Subheader 1_5">
</a>


Q5) Can you get these facts about “extent” or “size” in more than one way? Can you imagine different functions being useful in different contexts?
 
```{r}
#way 1 using dim command
dim(gapminder)[1] #Rows
dim(gapminder)[2] #Coloumns
#way two using length and transpose command
length(transpose(gapminder)) #Rows
length(gapminder)  #Coloumns
```


<a name="Subheader 1_6">
</a>

Q6) What data type is each variable?

```{r}
sapply(gapminder,typeof)
```

<a name="Header_2">
</a>

# Exploring individual variables

<a name="Subheader 2_1">
</a>

Q1) What are possible values (or range, whichever is appropriate) of each variable?

```{r}
max <- sapply(select(gapminder,year,lifeExp,pop,gdpPercap),max)
min <- sapply(select(gapminder,year,lifeExp,pop,gdpPercap),min)
#Range of Years
c(max[1],min[1])
#Range of LifeExp
c(max[2],min[2])
#Range of Pop
c(max[3],min[3])
#Range of gdpPercap
c(max[4],min[4])
table(gapminder$country)
table(gapminder$continent)
```

Another method using function range (simple!!!)

```{r}
sapply(select(gapminder,year,lifeExp,pop,gdpPercap),range)
```

Another method using summary (Yay!!!)

```{r}
summary(gapminder)
```


<a name="Subheader 2_2">
</a>


Q2) What values are typical? What’s the spread? What’s the distribution?

For findinding the distribution and typical values graphs are plotted, from that peak and spread can be visualised. Peak represent typical values. 

```{r}
#Exploring continents
ggplot(gapminder, aes(x = continent)) + geom_bar()
#Exploring distribution of population (spread and typical values)
#population based on entire glob
ggplot(gapminder, aes(x = log10(pop))) + geom_histogram()
#Polpulation contribution by each continent
ggplot(gapminder, aes(x = log10(pop), fill = continent )) + geom_histogram()
#Polpulation distribution in each continent
#Africa
gapminder %>% 
  filter(continent == "Africa") %>% 
  ggplot(aes(x = log10(pop))) + geom_histogram()
#America
gapminder %>% 
  filter(continent == "Americas") %>% 
  ggplot(aes(x = log10(pop))) + geom_histogram()
#Asia
gapminder %>% 
  filter(continent == "Asia") %>% 
  ggplot(aes(x = log10(pop))) + geom_histogram()
#Europe
gapminder %>% 
  filter(continent == "Europe") %>% 
  ggplot(aes(x = log10(pop))) + geom_histogram()
#Oceania
gapminder %>% 
  filter(continent == "Oceania") %>% 
  ggplot(aes(x = log10(pop))) + geom_histogram()
#Exploring distribution of lifeExp (spread and typical values)
#lifeExp based on entire glob
ggplot(gapminder, aes(x = lifeExp)) + geom_density()
#lifeExp contribution by each continent
ggplot(gapminder, aes(x = lifeExp, color = continent)) + geom_density()
#Exploring distribution of gdpPercap (spread and typical values)
#gdpPercap based on entire glob
ggplot(gapminder, aes(x = log10(gdpPercap))) + geom_density()
#gdpPercap ontribution by each continent
ggplot(gapminder, aes(x = log10(gdpPercap), color = continent)) + geom_density()
```

<a name="Header_3">
</a>

# Exploring various plot types

<a name="Subheader 3_1">
</a>


Q1) A scatterplot of two quantitative variables.
```{r}
gapminder %>% 
  filter(country %in% c("India","China","Canada","Australia","Japan")) %>% 
  mutate(gdp = gdpPercap * pop) %>% 
  ggplot(aes(x = log10(gdp), y = lifeExp)) +
  geom_point(aes(color = country))
gapminder %>% 
  group_by(country) %>% 
  summarise(mean_pop=log10(mean(pop)), mean_life=mean(lifeExp)) %>% 
  ggplot(aes(x=mean_pop, y=mean_life)) + geom_point()
```


<a name="Subheader 3_2">
</a>

Q2) A plot of one quantitative variable. Maybe a histogram or densityplot or frequency polygon.
```{r}
gapminder %>% 
  filter(country %in% c("India","China","Canada","Australia","Japan")) %>% 
  ggplot(aes(x = log10(pop), color=country)) + geom_histogram()
```


<a name="Subheader 3_3">
</a>

Q3) A plot of one quantitative variable and one categorical. Maybe boxplots for several continents or countries.
```{r}
gapminder %>% 
  filter(country %in% c("India","China","Canada","Australia","Japan")) %>% 
  ggplot(aes(x=country,y=log10(pop))) + geom_boxplot()
gapminder %>% 
  filter(country %in% c("Canada", "Australia")) %>% 
  ggplot(aes(x = country, y = pop)) + geom_jitter()
gapminder %>% 
  mutate(gdp = gdpPercap * pop) %>% 
  ggplot(aes(x=continent, y = log10(gdp))) + geom_boxplot()
```



