---
title: "Exploring Gapminder"
output: github_document
---

# Table of content
- Determining attributes of gapminder (Smell test the data) [Here](#Header_1)
- Exploring individual variables [Here](#Header_2)
- Exploring various plot types [Here](#Header_3)

<a name="Header_1">
</a>

# Determining attributes of gapminder (Smell test the data)


Loading the libraries

```{r,message=FALSE}
library(gapminder)
library(tidyverse)
library(ggplot2)
```

Q1) What type is gapminder?
```{r}
typeof(gapminder)
```

Q2) What is the class on gapminder

```{r}
class(gapminder)
```


Q3) How many variables/columns?

```{r}
ncol(gapminder)
```

Q4) How many rows/observations?

```{r}
nrow(gapminder)
```

Q5) Can you get these facts about “extent” or “size” in more than one way? Can you imagine different functions being useful in different contexts?
 
```{r}
#way 1 using dim command
dim(gapminder)[1] #Rows
dim(gapminder)[2] #Coloumns
#way two using length and transpose command
length(transpose(gapminder)) #Rows
length(gapminder)  #Coloumns
```

Q6) What data type is each variable?

```{r}
sapply(gapminder,typeof)
```

<a name="Header_2">
</a>

# Exploring individual variables

Q1) What are possible values (or range, whichever is appropriate) of each variable?

```{r}
max <- sapply(select(gapminder,year,lifeExp,pop,gdpPercap),max)
min <- sapply(select(gapminder,year,lifeExp,pop,gdpPercap),min)
#Range of Years
c(max[1],min[1])
#Range of LifeExp
c(max[2],min[2])
#Range of Pop
c(max[3],min[3])
#Range of gdpPercap
c(max[4],min[4])
table(gapminder$country)
table(gapminder$continent)
```

Another method using function range (simple!!!)

```{r}
sapply(select(gapminder,year,lifeExp,pop,gdpPercap),range)
```

Another method using summary (Yay!!!)

```{r}
summary(gapminder)
```

Q2) What values are typical? What’s the spread? What’s the distribution?
For findinding the distribution and typical values graphs are plotted, from that peak and spread can be visualised. Peak represent typical values. 

```{r}
#Exploring continents
ggplot(gapminder, aes(x = continent)) + geom_bar()
#Exploring distribution of population (spread and typical values)
#population based on entire glob
ggplot(gapminder, aes(x = log10(pop))) + geom_histogram()
#Polpulation contribution by each continent
ggplot(gapminder, aes(x = log10(pop), fill = continent )) + geom_histogram()
#Polpulation distribution in each continent
#Africa
gapminder %>% 
  filter(continent == "Africa") %>% 
  ggplot(aes(x = log10(pop))) + geom_histogram()
#America
gapminder %>% 
  filter(continent == "Americas") %>% 
  ggplot(aes(x = log10(pop))) + geom_histogram()
#Asia
gapminder %>% 
  filter(continent == "Asia") %>% 
  ggplot(aes(x = log10(pop))) + geom_histogram()
#Europe
gapminder %>% 
  filter(continent == "Europe") %>% 
  ggplot(aes(x = log10(pop))) + geom_histogram()
#Oceania
gapminder %>% 
  filter(continent == "Oceania") %>% 
  ggplot(aes(x = log10(pop))) + geom_histogram()
#Exploring distribution of lifeExp (spread and typical values)
#lifeExp based on entire glob
ggplot(gapminder, aes(x = lifeExp)) + geom_density()
#lifeExp contribution by each continent
ggplot(gapminder, aes(x = lifeExp, color = continent)) + geom_density()
#Exploring distribution of gdpPercap (spread and typical values)
#gdpPercap based on entire glob
ggplot(gapminder, aes(x = log10(gdpPercap))) + geom_density()
#gdpPercap ontribution by each continent
ggplot(gapminder, aes(x = log10(gdpPercap), color = continent)) + geom_density()
```

<a name="Header_3">
</a>

# Exploring various plot types


```{r}
gapminder %>% 
  filter(country %in% c("Canada", "Australia")) %>% 
  mutate(gdp = gdpPercap * pop) %>% 
  ggplot(aes(x = log10(gdp), y = lifeExp)) +
  geom_point(aes(color = country))
gapminder %>% 
  filter(country %in% c("Canada", "Australia")) %>% 
  ggplot(aes(x = country, y = pop)) + geom_jitter()
gapminder %>% 
  filter(country %in% c("Canada")) %>% 
  ggplot(aes(x = gdpPercap)) + geom_histogram()
  
#ggplot(gapminder, aes(x = log10(gdpPercap), y = lifeExp)) +
#  geom_point(aes(color = continent))
#ggplot(gapminder, aes(x = continent, y = lifeExp)) + geom_jitter()+geom_boxplot()
```



