---
title: "Hw05"
output: github_document
---

# Table of Content

- Factor management
    - Factorise. Transform some of the variable in the singer_locations dataframe into factors: pay attention at what levels you introduce and their order. Try and consider the difference between the base R as.factor and the forcats-provided functions [here](#header1)
    - Drop Oceania. Filter the Gapminder data to remove observations associated with the continent of Oceania. Additionally, remove unused factor levels. Provide concrete information on the data before and after removing these rows and Oceania; address the number of rows and the levels of the affected factors [here](#header2)
    - Reorder the levels of country or continent. Use the forcats package to change the order of the factor levels, based on a principled summary of one of the quantitative variables. Consider experimenting with a summary statistic beyond the most basic choice of the median [here](#header3)
- n

Just to show the difference between as.factor() and forcats as_factor() I have considered using singer_location data but for all the rest I have used the gapminder data itself.


```{r, message=FALSE}
library(singer)
library(gapminder)
library(tidyverse)
library(forcats)
```

## Factor management

<a name=header1>
</a>

### Factorise. Transform some of the variable in the singer_locations dataframe into factors: pay attention at what levels you introduce and their order. Try and consider the difference between the base R as.factor and the forcats-provided functions

```{r}
new_set <- singer_locations %>% mutate(title_forcats_factor = as_factor(title), title_base_factor = as.factor(title))
knitr::kable(head(levels(new_set$title_forcats_factor),10))
knitr::kable(head(levels(new_set$title_base_factor),10))
knitr::kable(head(new_set$title,10))
```

<a name=header2>
</a>

### Drop Oceania. Filter the Gapminder data to remove observations associated with the continent of Oceania. Additionally, remove unused factor levels. Provide concrete information on the data before and after removing these rows and Oceania; address the number of rows and the levels of the affected factors.

```{r}
nrow(gapminder)
length(levels(gapminder$continent))
new_set <-gapminder %>% filter(continent != "Oceania") %>% droplevels()
new_set_different <- gapminder %>% filter(continent != "Oceania")
nrow(new_set)
length(levels(new_set$continent))
length(levels(new_set_different$continent))
```


<a name=header2>
</a>

### Reorder the levels of country or continent. Use the forcats package to change the order of the factor levels, based on a principled summary of one of the quantitative variables. Consider experimenting with a summary statistic beyond the most basic choice of the median.

```{r}
ggplot(new_set,aes(continent,pop)) + geom_boxplot() + scale_y_log10()
new_set %>%  arrange(pop) %>% ggplot(aes(continent,pop)) + geom_boxplot() + scale_y_log10()
ggplot(new_set,aes(fct_reorder(continent,pop),pop)) + geom_boxplot() + scale_y_log10()
new_set %>% mutate(continent = fct_reorder(continent,pop,mean)) %>% ggplot(aes(continent,pop)) + geom_boxplot() + scale_y_log10() 
new_set %>% mutate(continent = fct_reorder(continent,pop,mean)) %>% arrange(pop) %>% ggplot(aes(continent,pop)) + geom_boxplot() + scale_y_log10()

new_set %>% mutate(continent = fct_reorder(continent,pop,mean)) %>% select(continent) %>% write_csv("continent.csv")

new_set %>% mutate(continent = fct_reorder(continent,pop,mean)) %>% select(continent) %>% saveRDS("continent.rds")

new_set %>% mutate(continent = fct_reorder(continent,pop,mean)) %>% select(continent) %>% dput("continent.txt")

continen_csv <- read_csv("continent.csv")
continen_RDS <- readRDS("continent.rds")   
continen_txt <- dget("continent.txt")  

comp <- tibble(level_original = head(levels(fct_reorder(new_set$continent,new_set$pop,mean))))
comp
continen_csv <- continen_csv %>% mutate(continent = factor(continent))
continen_RDS <- continen_RDS %>% mutate(continent = factor(continent))
continen_txt <- continen_txt %>% mutate(continent = factor(continent))
comp %>% mutate(level_csv = head(levels(continen_csv$continent)),level_RDS = head(levels(continen_RDS$continent)),level_txt = head(levels(continen_txt$continent)))


```




